---
title: "Carly's Final Data Science Presentation"
author: "Carly Baretz"
subtitle: "December 6, 2024"
format:
  revealjs:
    scrollable: true
    slide-number: true
    show-slide-number: all
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

## Salary Comparison of Pitchers: New York Mets vs. New York Yankees (2010)

This analysis investigates salary differences between pitchers from the New York Mets and New York Yankees in the 2010 MLB season


MLB data from the openintro package (2010 season)

## MLB Salaries 

```{r}
#| echo = FALSE  

# Load necessary libraries
library(tidyverse)
library(openintro)

# Load the MLB dataset
data("mlb")

# Data preparation: Calculate average salary by team
team_salaries <- mlb |>
  group_by(team) |>
  summarize(avg_salary = mean(salary, na.rm = TRUE)) |>
  arrange(desc(avg_salary))

# Bar chart: Average salary by team
ggplot(team_salaries, aes(x = reorder(team, avg_salary), y = avg_salary)) +
  geom_bar(stat = "identity", fill = "blue") +
  coord_flip() +
  labs(
    title = "Average MLB Salaries by Team (2010)",
    x = "Team",
    y = "Average Salary (in thousands of dollars)"
  ) 

```

## Mets and Yankees Pitcher Salaries 

```{r}

mlb_pitchers <- mlb |>
  filter((team == "New York Mets" | team == "New York Yankees") & position == "Pitcher" & !is.na(salary)) |>
  select(team, salary)
mlb_pitchers
```

\* Salaries are represented in the thousands, so 20,144 is actually $20,144,000. 

## Observed Difference
```{r}
# Calculate the observed difference in average salary between Mets and Yankees
obs_diff <- mlb_pitchers |>
  group_by(team) |>
  summarize(obs_ave = mean(salary)) |>
  summarize(obs_ave_diff = diff(obs_ave)) |>
  pull(obs_ave_diff)

# Check if obs_diff was calculated correctly
obs_diff

```

## Permutation Test
```{r}
perm_data <- function(rep, data) {
  data |>
    mutate(salary_perm = sample(salary, replace = FALSE)) |>
    group_by(team) |>
    summarize(perm_ave = mean(salary_perm)) |>
    summarize(perm_ave_diff = diff(perm_ave), rep = rep)
}
set.seed(47)

perm_stats <- map(1:500, perm_data, data = mlb_pitchers) |>
  list_rbind()

perm_stats
```

## Visualization

```{r}
#| echo = FALSE 

# Plot the permutation distribution with the observed difference
ggplot(perm_stats, aes(x = perm_ave_diff)) +
  geom_histogram(bins = 50, fill = "lightblue", color = "black") +
  geom_vline(xintercept = obs_diff, color = "red", linetype = "dashed") +
  labs(
    title = "Permutation Test: Difference in Mean Salary (Mets vs. Yankees Pitchers)",
    x = "Difference in Permuted Mean Salary",
    y = "Frequency",
    caption = "Source: OpenIntro MLB Dataset"
  ) 
```
## Conclusion 

Pitchers on the Mets earned a higher average salary than Yankees pitchers in 2010. 

HOWEVER, the difference is not statistically significant. 
