---
title: "Netflix Data Analysis"
description: |
author: Carly Baretz 
date: October 2, 2024
format: html

execute:
  warning: false
  message: false
---

Netflix is the biggest streaming service in the world, offering a wide range of content in various genres. This analysis examines Netflix's catalog using data from the Tidy Tuesday project (April 20, 2021). I will explore the most prevalent genres on Netflix, patterns in the word count of titles and their use of common terms like "Love," "Story," or "Adventure,"and trends in the growth of Netflix's catalog over time.

```{r}
#| echo: false 

library(tidyverse)
library(ggplot2)
library(dplyr)
library(stringr)
# Suppress warnings and messages
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
netflix_titles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-20/netflix_titles.csv')

# Calculate the number of genres listed for each title
netflix_titles <- netflix_titles |>
  mutate(genre_count = str_count(listed_in, ",") + 1)

# Separate genres into rows and calculate counts
genres <- netflix_titles |>
  separate_rows(listed_in, sep = ",")|>
  group_by(listed_in) |>
  summarise(count = n()) |>
  arrange(desc(count))

# Plot the top 10 genres by count
ggplot(genres[1:10,], aes(x = reorder(listed_in, -count), y = count)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Top 10 Genres by Number of Titles",
       x = "Genre", y = "Count") +
  theme_minimal() +
  coord_flip()

```

The bar graph above shows the top 10 genres on Netflix by amount of titles. You can see that Netflix's largest genre by which is has the most titles is international movies, with Netflix having upwards of 2,500 international movies.

```{r}

netflix_titles <- netflix_titles |>
  mutate(
    title_word_count = str_count(title, "\\b\\w+\\b"),
    contains_common_word = str_detect(title, "\\b(Love|Story|Adventure)\\b")
  )

title_word_distribution <- netflix_titles |>
  group_by(title_word_count, contains_common_word) |>
  summarise(count = n()) |>
  arrange(desc(count))


ggplot(title_word_distribution, aes(x = factor(title_word_count), y = count, fill = contains_common_word)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Distribution of Netflix Titles by Word Count and Common Words (Love, Story, Adventure)",
       x = "Number of Words in Title", y = "Count of Titles",
       fill = "Contains 'Love', 'Story', or 'Adventure'") +
  theme_minimal()

```

This graph shows the distribution of number of words in titles on Netflix, showing that most of the titles on Netflix are 2 words long. Additionally, the graph shows the prevalence of titles with 3 common words: love, story, and adventure. The graph shows that most of the titles do not contain "love," "story," or "adventure," but it shows that if the words do appear, they show up in titles ranging from 1 word long to 13 words long.

```{r}

netflix_titles <- netflix_titles |>
  mutate(year_added = str_extract(date_added, "\\d{4}"))

netflix_titles <- netflix_titles |>
  filter(!is.na(year_added))

titles_by_year <- netflix_titles |>
  group_by(year_added) |>
  summarise(count = n()) |>
  arrange(year_added)

ggplot(titles_by_year, aes(x = year_added, y = count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Number of Netflix Titles Added by Year",
       x = "Year", y = "Number of Titles") +
  theme_minimal()

```

This graph shows the number of titles of shows or movies added to Netflix each year. The first titles were added to Netflix in 2008 and in 2019 the highest number of titles were added to Netflix.

To look for patterns in the dataset to then visualize, I started by identifying how many Netflix titles have a year in the title name, have a colon in the title, or start with "The."

```{r}
# Add columns based on regular expression patterns
netflix_titles <- netflix_titles |>
  mutate(
    contains_year = str_detect(title, "\\b(19|20)\\d{2}\\b"),  # Detect years like 1999 or 2021
    contains_colon = str_detect(title, ":"),                  # Detect titles with colons
    starts_with_the = str_detect(title, "^The\\b")            # Detect titles starting with "The"
  )

# Summarize the counts of titles matching each pattern
pattern_summary <- netflix_titles |>
  summarize(
    with_year = sum(contains_year, na.rm = TRUE),
    with_colon = sum(contains_colon, na.rm = TRUE),
    starts_with_the = sum(starts_with_the, na.rm = TRUE)
  )

pattern_summary

```

Next I made a bar plot to visualize the amount of Netflix titles with the given pattern.

```{r}

#| echo: false 
#| eval: true 
# Prepare data for visualization
pattern_data <- pattern_summary |>
  pivot_longer(cols = everything(), names_to = "Pattern", values_to = "Count") |>
  mutate(Pattern = recode(Pattern,
                          "with_year" = "Contains Year",
                          "with_colon" = "Contains Colon",
                          "starts_with_the" = "Starts with 'The'"))

# Plot the counts of each pattern
ggplot(pattern_data, aes(x = reorder(Pattern, Count), y = Count, fill = Pattern)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Frequency of Specific Patterns in Netflix Titles",
    x = "Pattern",
    y = "Count",
    fill = "Pattern",
    caption = "Source: Netflix Movies and TV Shows Dataset (Tidy Tuesday)"
  ) +
  theme_minimal()

```

Finally, I made a bar plot to visualize the distribution of years mentioned in Netflix titles. Below is the spread of Netflix titles containing years, and the amount of titles with names a year.

```{r}

# Extract titles with years and summarize their counts
titles_with_years <- netflix_titles |>
  filter(contains_year) |>
  mutate(year_in_title = str_extract(title, "\\b(19|20)\\d{2}\\b")) |>
  group_by(year_in_title) |>
  summarize(count = n()) |>
  arrange(desc(count))

# Plot the distribution of years in Netflix titles
ggplot(titles_with_years, aes(x = year_in_title, y = count)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(
    title = "Distribution of Years Mentioned in Netflix Titles",
    x = "Year in Title",
    y = "Count",
    caption = "Source: Netflix Movies and TV Shows Dataset (Tidy Tuesday)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

The dataset was sourced from the Tidy Tuesday project (April 20, 2021) and is originally from Kaggle's Netflix Movies and TV Shows dataset. The dataset aggregates publicly available information on Netflix's catalog, including title, genre, date added, director, cast, and release year.
