---
title: "Analysis of Scientific Research Data"
description: |
date: November 21, 2024
format: html

execute:
  warning: false
  message: false
---

This project explores data from the Wideband Absorbance Investigation (WAI) database to analyze the relationship between frequency and absorbance in various studies. Using SQL (Structured Query Language), I queried the database to extract and manipulate data, such as calculating average absorbance values at specific frequencies and grouping data by study identifiers. 

SQL is a powerful language used for managing and analyzing large datasets stored in relational databases. After extracting the data, I used R to visualize trends through graphs, including comparisons between male and female participants in a specific study. This project highlights how database analysis and visualization can uncover patterns in scientific research data.


```{r}
#| echo = FALSE
library(mdsr)
library(dbplyr)
library(DBI)
library(RMariaDB)
library(tidyverse)
```

```{r}


con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)
Measurements <- tbl(con_wai, "Measurements")
PI_Info <- tbl(con_wai, "PI_Info")
Subjects <- tbl(con_wai, "Subjects")



```

To understand the data structure, I explored the WAI database by listing all available tables. 

```{sql}
#| connection: con_wai
SHOW TABLES;
```

Next I viewed the first few rows of the "Measurements" table, which allows us to understand the variables and structure of the dataset. 

```{sql}
#| connection: con_wai
SELECT *
FROM Measurements
LIMIT 0, 5;
```

Next I only selected information about a specific study. 

```{sql}
#| connection: con_wai
SELECT * FROM PI_Info 
WHERE Identifier = "Abur_2014"; 
```

Next I identified data from the "Measurements" table about a specific study. 

```{sql}
#| connection: con_wai
SELECT * FROM Measurements 
WHERE Identifier = "Abur_2014"; 
```

Next I selected data from the Measurements table, selecting specific Identifiers and filtering rows where Frequency is between 200 and 8000. Then I calculated the base-10 logarithm of Frequency (log_frequency) and the average absorbance (mean_absorbance), and grouped results by Identifier and Frequency.

```{sql}
#| connection: con_wai
#| output.var: "graph"

SELECT Identifier, Frequency, LOG10(Frequency) AS log_frequency,  AVG(Absorbance) AS mean_absorbance 
FROM Measurements
WHERE Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008"
"Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000
GROUP BY Identifier, Frequency;  


```

The plot visualizes how absorbance changes with frequency for given studies
```{r}
graph |>
ggplot(aes (x=Frequency, y = mean_absorbance, 
   color =  Identifier,
   group = Identifier)) +
  geom_line()+
  scale_x_log10()+
      labs(
    title = "Mean Absorbance From Each Publication in the WAI Database",   
    x = "Frequency (Hz)",           
    y = "Mean Absorbance"  )

```

This table links multiple tables in the WAI database by counting distinct ears measured in the study with measurement data. 

```{sql}
#| connection: con_wai
 SELECT p.Identifier, p.Year, p.AuthorsShortList, 
 COUNT(DISTINCT SubjectNumber, Ear) AS ear_you 
FROM PI_Info AS p 
LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier 
WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008"
"Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000
GROUP BY Identifier, Instrument;
```
This creates 1 table to combine data from different tables. 
```{sql}
#| connection: con_wai
 SELECT p.Identifier, Year, AuthorsShortList, 
 COUNT(DISTINCT SubjectNumber, Ear) AS ear_you,
 CONCAT(AuthorsShortList, " (" , year, ") " , "N=" , 
 COUNT(DISTINCT SubjectNumber, Ear), "; ", Instrument) AS legend 
FROM PI_Info AS p 
LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier 
WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008"
"Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000
GROUP BY Identifier, Instrument;

```

This adds on to the table above, including mean absorbance across frequencies. 

```{sql}
#| connection: con_wai
#| output.var: "graph2"
 SELECT p.Identifier, p.Year, p.AuthorsShortList, Frequency,
 LOG10(Frequency) AS log_frequency, AVG(Absorbance) AS mean_absorbance,
 COUNT(DISTINCT SubjectNumber, Ear) AS ear_you,
 CONCAT(AuthorsShortList, " (" , year, ") ", "N=", COUNT(DISTINCT SubjectNumber, Ear), "; ", Instrument) AS legend
FROM PI_Info AS p 
LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier 
WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008"
"Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000
GROUP BY Identifier, Instrument, Frequency;
```

```{r}
graph2 |>
ggplot(aes (x=Frequency, y = mean_absorbance, 
   color =  legend,
   group = legend)) +
  geom_line()+
  scale_x_log10()+
        labs(
    title = "Mean Absorbance From Each Publication in the WAI Database",
    legend = "",
    x = "Frequency (Hz)",           
    y = "Mean Absorbance",
    color = NULL)


```

This looks at all the information in the "Subjects" table. 

```{sql}
#| connection: con_wai
SELECT * 
FROM Subjects

```

This calculates mean absorbance for males and females separately in the Aithal 2013 study. 
```{sql}
#| connection: con_wai
#| output.var: "graphforQ2"
SELECT Sex, Frequency, AVG(Absorbance) AS mean_absorbance, m.Identifier 
FROM Subjects AS s 
RIGHT JOIN Measurements AS m ON m.SubjectNumber = s.SubjectNumber
WHERE m.Identifier = "Aithal_2013" AND Frequency > 200 AND Frequency < 8000
GROUP BY m.Identifier, Frequency, Sex;

```

```{r}
head(graphforQ2)
```
This visualizes absorbance trends by gender in the Aithal 2013 study. 

```{r}
graphforQ2 |>
  ggplot(aes(
    x = Frequency,
    y = mean_absorbance,
    color = Sex,
    group = Sex
  )) +
  geom_line() +
  scale_x_log10() +
  labs(
    title = "Mean Absorbance Across Frequencies by Gender in 'Aithal_2013' Study", 
    x = "Frequency",
    y = "Mean Absorbance", 
    color = "Gender"  
  ) +
  facet_wrap(~Sex)  
```

In this project, I analyzed data from the WAI database to explore the relationship between frequency and absorbance across various studies. I used SQL queries to extract, filter, and calculate the average absorbance values for specific frequencies and grouped the data by study identifiers. These results were then visualized in R using ggplot, with plots on a logarithmic frequency scale to highlight trends. Additionally, I compared absorbance data by sex for a specific study to observe any potential differences, finding that male and female values were the same.

I got my data from the WAI Database and the graph is modeled from a graph in this study: Voss, Susan E. Ph.D. Resource Review. Ear and Hearing 40(6):p 1481, November/December 2019. | DOI: 10.1097/AUD.0000000000000790 
