---
title: "Project 4"
description: |
author: Carly Baretz 
date: November 21, 2024
format: html

execute:
  warning: false
  message: false
---

To complete this project, I plan to look at the WAI database and select the tables which 
detail the studies and information about the studies. I then plan on finding the average 
absorbence of the data and plot that against the frequency of the data. 

```{r}
library(mdsr)
library(dbplyr)
library(DBI)
```

```{r}
library(RMariaDB)
library(tidyverse)

con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)
Measurements <- tbl(con_wai, "Measurements")
PI_Info <- tbl(con_wai, "PI_Info")
Subjects <- tbl(con_wai, "Subjects")

# collect(Measurements)

```

```{sql}
#| connection: con_wai
SHOW TABLES;
```

This table shows all the available tables in the WAI database. 

```{sql}
#| connection: con_wai
SELECT *
FROM Measurements
LIMIT 0, 5;
```
This table shows what the measurements table looks like and the information it holds.

```{sql}
#| connection: con_wai
SELECT *
FROM PI_Info
LIMIT 0, 50;
```
This is the PI Info table. 

```{sql}
#| connection: con_wai
SELECT * FROM PI_Info 
WHERE Identifier = "Abur_2014"; 
```
This is one specific row from the PI Info table and it gives details about one specific study
```{sql}
#| connection: con_wai
SELECT * FROM Measurements 
WHERE Identifier = "Abur_2014"; 
```
This is more information about the study.

```{sql}
#| connection: con_wai
#| output.var: "graph"

SELECT Identifier, Frequency, LOG10(Frequency) AS log_frequency,  AVG(Absorbance) AS mean_absorbance 
FROM Measurements
WHERE Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008"
"Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000
GROUP BY Identifier, Frequency;  


```

```{r}
graph |>
ggplot(aes (x=Frequency, y = mean_absorbance, 
   color =  Identifier,
   group = Identifier)) +
  geom_line()+
  scale_x_log10()+
      labs(
    title = "Mean Absorbance From Each Publication in the WAI Database",   
    x = "Frequency (Hz)",           
    y = "Mean Absorbance"  )


```
This graph shows the mean absorbance From each publication in the WAI database
```{sql}
#| connection: con_wai
 SELECT p.Identifier, p.Year, p.AuthorsShortList, 
 COUNT(DISTINCT SubjectNumber, Ear) AS ear_you 
FROM PI_Info AS p 
LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier 
WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008"
"Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000
GROUP BY Identifier, Instrument;
```
This table shows information about the publications in the database. 

```{sql}
#| connection: con_wai
 SELECT p.Identifier, Year, AuthorsShortList, 
 COUNT(DISTINCT SubjectNumber, Ear) AS ears,
 CONCAT(AuthorsShortList, " (" , year, ") " , "N=" , 
 COUNT(DISTINCT SubjectNumber, Ear), "; ", Instrument) AS Legend 
FROM PI_Info AS p 
LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier 
WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008"
"Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000
GROUP BY Identifier, Instrument;

```
This table shows information about the publications in the database. 

```{sql}
#| connection: con_wai
#| output.var: "graph2"
 SELECT p.Identifier, p.Year, p.AuthorsShortList, Frequency,
 LOG10(Frequency) AS log_frequency, AVG(Absorbance) AS mean_absorbance,
 COUNT(DISTINCT SubjectNumber, Ear) AS ear_u,
 CONCAT(AuthorsShortList, " (" , year, ") ", "N=", COUNT(DISTINCT SubjectNumber, Ear), "; ", Instrument) AS legend
FROM PI_Info AS p 
LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier 
WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008"
"Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000
GROUP BY Identifier, Instrument, Frequency;
```

```{r}
graph2 |>
ggplot(aes (x=Frequency, y = mean_absorbance, 
   color =  legend,
   group = legend)) +
  geom_line()+
  scale_x_log10()+
        labs(
    title = "Mean Absorbance From Each Publication in the WAI Database",   
    x = "Frequency (Hz)",           
    y = "Mean Absorbance"  )


```
This graph shows the mean absorbance From each publication in the WAI database, 
with details about measurement and equipement used in each study. 
```{sql}
#| connection: con_wai
SELECT * 
FROM Subjects

```
This table shows the data in the subjects table. 
```{sql}
#| connection: con_wai
#| output.var: "graphforQ2"
SELECT Sex, Frequency, AVG(Absorbance) AS mean_absorbance, m.Identifier 
FROM Subjects AS s 
RIGHT JOIN Measurements AS m ON m.Identifier = s.Identifier
WHERE m.Identifier = "Aithal_2013" AND Frequency > 200 AND Frequency < 8000
GROUP BY m.Identifier, Frequency, Sex;

```

```{r}
head(graphforQ2)
```
This tabe shoes the frequency, mean absorbance, and sex of participants for one 
specific study, Aithal 2013. 
```{r}
graphforQ2 |>
  filter(Sex == "Female") |>
ggplot(aes (x=Frequency, y = mean_absorbance, 
   color =  Sex,
   group = Sex)) +
  geom_line() +
  scale_x_log10()+
    labs(
    title = "Mean Absorbance vs Frequency for Women",   
    x = "Frequency",           
    y = "Mean Absorbance"  )


```
This graph shows the frequency and absorbance for women in the study. 

```{r}
graphforQ2 |>
  filter(Sex == "Male") |>
ggplot(aes (x=Frequency, y = mean_absorbance, 
   color =  Sex,
   group = Sex)) +
  geom_line() +
  scale_x_log10()+
  labs(
    title = "Mean Absorbance vs Frequency for Men",   
    x = "Frequency",           
    y = "Mean Absorbance"  )

```
This graph shows the frequency and absorbance for men in the study. 

In this project, I analyzed data from the WAI database to explore the relationship between frequency and absorbance across various studies. I used SQL queries to extract, filter, and calculate the average absorbance values for specific frequencies and grouped the data by study identifiers. These results were then visualized in R using ggplot, with plots on a logarithmic frequency scale to highlight trends. Additionally, I compared absorbance data by sex for a specific study to observe any potential differences, finding that male and female values were the same.


I got my data from the WAI Database and the graph is modeled from a graph in this study: https://pmc.ncbi.nlm.nih.gov/articles/PMC7093226/
